On affine extension of splint root systems

### 2013-03-04 AffineLieAlgebras/Splints/affine.tex
Some text on affine splints

>>>>

\section{Splints and affine root systems}
\label{sec:splints-affine-root}

\begin{Def}
  Embedding $\phi$ of a root system $\Delta_1$ into a root system $\Delta$ is a bijective map of
  roots of $\Delta_{1}$ to a (proper) subset of $\Delta$ that commutes with vector composition law
  in $\Delta_{1}$ and $\Delta$.
\begin{equation*}
\phi:\Delta_1 \longrightarrow \Delta, \quad \phi \circ (\alpha + \beta) =\phi \circ \alpha + \phi \circ \beta,\,\,\, \alpha,\beta \in \Delta_1
\end{equation*}
\end{Def}

Note that the image $Im(\phi)$ must not inherit the root system properties except the addition rules
equivalent to the addition rules in $\Delta_{1}$ (for pre-images). Two embeddings $\phi_1$ and
$\phi_2$ can splinter $\Delta$ when the latter can be presented as a disjoint union of images
$Im(\phi_1)$ and $Im(\phi_2)$.

$\phi$ induces an injection of formal algebras $:{\mathcal{E}}_0
\hookrightarrow \mathcal{E}$ and for the image ${\mathcal{E}}%
_i=Im_{\phi}\left( {\mathcal{E}}_0\right)$ one can consider its inverse $%
\phi^{-1}:{\mathcal{E}}_i \longrightarrow {\mathcal{E}}_0$.

\begin{Def}
  A root system $\Delta $ ''splinters'' as $(\Delta _{1},\Delta _{2})$ if there are two embeddings
  $\phi _{1}:\Delta _{1}\hookrightarrow \Delta $ and $%
  \phi _{2}:\Delta _{2}\hookrightarrow \Delta $ where (a) $\Delta $ is the disjoint union of the
  images of $\phi _{1}$ and $\phi _{2}$ and (b) neither the rank of $\Delta _{1}$ nor the rank of
  $\Delta _{2}$ exceeds the rank of $%
  \Delta $.
\end{Def}

It is equivalent to say that $(\Delta_1,\Delta_2)$ is a "splint'' of $\Delta$ and we shall denote
this by $\Delta \approx (\Delta_1,\Delta_2)$. Each component $\Delta_1$ and $\Delta_2$ is a "stem''
of the splint.

We consider the case when one of the stems $\Delta _{1}=\Delta _{\frak{a}}$ is a root subsystem. As shown in paper \cite{2011arXiv1111.6787L} the second stem $\Delta _{\frak{s}}:=\Delta
_{2}=\Delta \setminus \Delta _{\frak{a}}$ can be translated into a product
$\prod_{\beta \in \Delta _{\frak{s}}^{+}}\left( 1-e^{-\beta }\right)=-\sum_{\gamma \in P}s(\gamma )e^{-\gamma }\quad   \label{splint product}$
and it defines an injection fan $\Gamma _{\frak{a}%
\hookrightarrow \frak{g}}$ \cite{lyakhovsky1996rra,ilyin812pbc,2010arXiv1007.0318L}.

Using straightforward calculation we can easily obtain an identity on theta-functions

definition of Dedekind eta-function
$\eta(\tau)=q^{1/24}\prod_{n=1}^{\infty}(1-q^{n})$, where
$q=e^{2\pi i \tau}$ we can rewrite this identity as the relation
imposed on theta-functions $\Theta^{(\gfh)}_{\widehat{\lambda}=(\lambda,k,0)}(\tau,z)=\sum_{\xi\in Q_{\gf}+\frac{\lambda}{k}}e^{2\pi i k \left(\frac{1}{2} (\xi,\xi) \tau + (\xi,z)\right)}$:
\begin{multline}
  \label{eq:5}
  \eta(\tau)^{\mathrm{dim}(\af)}\prod_{\alpha\in\Delta_{1}^{+}}\frac{\Theta^{(\hat A_{1})}_{\alpha}(\tau,z)}{\eta(\tau)} \eta(\tau)^{\mathrm{dim}(\sfr)}\prod_{\beta\in\Delta_{2}^{+}}\frac{\Theta^{(\hat A_{1})}_{\phi\circ \beta}(\tau,z))}{\eta(\tau)}=\\
\eta(\tau)^{\mathrm{rank}(\af)+\mathrm{rank}(\sfr)-\mathrm{rank}(\gf)}
\eta(\tau)^{\mathrm{dim}(\gf)}\prod_{\alpha\in\Delta^{+}}\frac{\Theta^{(\hat A_{1})}_{\alpha}(\tau,z))}{\eta(\tau)}
\end{multline}
Here $z\in P_{\geq 0}\otimes \mathbb{C}$. Using Weyl denominator identity this relation can be
rewritten as a non-trivial relation connecting theta-functions of algebras $\gfh,\hat\sfr,\afh$:
\begin{equation}
  \label{eq:6}
  \left(\sum_{v\in W_{\af}}\epsilon(v) \Theta^{(\afh)}_{v\rho_{\af}}(\tau,z)\right)
  \cdot \left(\sum_{u\in W_{\sfr}}\epsilon(u) \Theta^{(\hat{\sfr})}_{\phi\circ(u\rho_{\sfr})}(\tau,z)\right)= 
  \left(\sum_{w\in W}\epsilon(w) \Theta^{(\gfh)}_{w\rho_{\gf}}(\tau,z)\right)
\end{equation}
